Combinatorial Proofs and Decomposition Theorems for First-order Logic

### 2021-01-24 main.tex
moved example
Add an example and a reference
moved big figs

\begin{equation*}
\vlderivation{
  \vlin{\conr}{}{\sqn{\exists x.(\cneg{p}x \vlor (\forall y.py))}}{
    \vlin{\exists}{}{\sqn{\exists x.(\cneg{p}x \vlor (\forall y.py)),\exists
x.(\cneg{p}x \vlor (\forall y.py))}}{
      \vlin{\vlor}{}{\sqn{\cneg{p}w \vlor (\forall y.py), \exists
x.(\cneg{p}x \vlor (\forall y.p(y)))}}{
        \vlin{\forall}{}{\sqn{\cneg{p}w, \forall y.py, \exists
x.(\cneg{p}x \vlor (\forall y.py))}}{
          \vlin{\exists}{}{\sqn{\cneg{p}w, pz, \exists
x.(\cneg{p}x \vlor (\forall y.py))}}{
            \vlin{\vlor}{}{\sqn{\cneg{p}w, pz, \cneg{p}z \vlor (\forall y.py))}}{
              \vlin{\weakr}{}{\sqn{\cneg{p}w, pz, \cneg{p}z, (\forall y.py))}}{
                \vlin{\weakr}{}{\sqn{\cneg{p}w, pz, \cneg{p}z}}{
                \vlin{\axr}{}{\sqn{pz, \cneg{p}z}}{
                  \vlhy{}}}}}}}}}}}
\end{equation*}
\juihsuan{LK1 and KS1 proofs of Peirce's law and LK1 proof of drinker formula}

@@

\figcps Fig.\,\ref{fig:cps} is a floating figure with four combinatorial proofs.

\figcpscondensed Fig.\,\ref{fig:cps-condensed} is a floating figure with the condensed forms of the four combinatorial proofs in Fig.\,\ref{fig:cps}.

@@

\begin{equation*}
\vlderivation{
  \vlin{\acD}{}{(\exists x. \cneg{p}x) \vlor (\forall y.(py \vlan pfy))}{
    \vlin{\mexD}{}{(\exists x. (\cneg{p}x \vlor \cneg{p}x) \vlor (\forall
y.(py \vlan pfy))}{
      \vlin{\fequ}{}{((\exists x. \cneg{p}x) \vlor (\exists x.\cneg{p}x))
\vlor (\forall y.(py \vlan pfy)}{
        \vlin{\exists}{}{\forall y.(((\exists x. \cneg{p}x) \vlor (\exists
x.\cneg{p}x)) \vlor (py \vlan pfy))}{
          \vlin{\exists}{}{\forall y.((\cneg{p}y \vlor (\exists x.\cneg{p}x))
\vlor (py \vlan pfy))}{
            \vlin{\fequ}{}{\forall y.((\cneg{p}y \vlor \cneg{p}fy) \vlor
(py \vlan pfy))}{
              \vlin{\sw}{}{\forall y.(\cneg{p}y \vlor ((py \vlan pfy)
\vlor \cneg{p}fy))}{
                \vlin{\fequ}{}{\forall y.(\cneg{p}y \vlor (py \vlan (pfy
\vlor \cneg{p}fy)))}{
                  \vlin{\aiD}{}{\forall y.((\cneg{p}y \vlor py) \vlan (pfy
\vlor \cneg{p}fy))}{
                    \vlin{\aiD}{}{\forall y.((\cneg{p}y \vlor py) \vlan
\ttt)}{
                      \vlin{\ttt}{}{\forall y.(\ttt \vlan \ttt)}{
                        \vlin{\forall}{}{\forall y.\ttt}{
                          \vlhy{\ttt}}}}}}}}}}}}}}
\end{equation*}
\juihsuan{the decomposed KS1 proof of $(\exists x. \cneg{p}(x)) \vlor (\forall
y.(p(y) \vlan p(f(y))))$}

@@

@@

@@

\begin{remark}
  In or definition of FOCP, we are slightly laxer than the original
  definition of~\cite{hughes:fopws}, as we allow for a variable
  renaming $\substof\phi$ which was forced to be the identity
  in~\cite{hughes:fopws}.
  %The reason for this relaxation in our case is that we want to cover a richer class of proofs, also allowing 
\end{remark}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

>>>>

\begin{equation*}
\vlderivation{
  \vlin{\conr}{}{\sqn{\exists x.(\cneg{p}x \vlor (\forall y.py))}}{
    \vlin{\exists}{}{\sqn{\exists x.(\cneg{p}x \vlor (\forall y.py)),\exists
x.(\cneg{p}x \vlor (\forall y.py))}}{
      \vlin{\vlor}{}{\sqn{\cneg{p}w \vlor (\forall y.py), \exists
x.(\cneg{p}x \vlor (\forall y.p(y)))}}{
        \vlin{\forall}{}{\sqn{\cneg{p}w, \forall y.py, \exists
x.(\cneg{p}x \vlor (\forall y.py))}}{
          \vlin{\exists}{}{\sqn{\cneg{p}w, pz, \exists
x.(\cneg{p}x \vlor (\forall y.py))}}{
            \vlin{\vlor}{}{\sqn{\cneg{p}w, pz, \cneg{p}z \vlor (\forall y.py))}}{
              \vlin{\weakr}{}{\sqn{\cneg{p}w, pz, \cneg{p}z, (\forall y.py))}}{
                \vlin{\weakr}{}{\sqn{\cneg{p}w, pz, \cneg{p}z}}{
                \vlin{\axr}{}{\sqn{pz, \cneg{p}z}}{
                  \vlhy{}}}}}}}}}}}
\end{equation*}
\juihsuan{LK1 and KS1 proofs of Peirce's law and LK1 proof of drinker formula}
\lutz{can you also do the KS1 for the drinker?}

@@

%%\figcps Fig.\,\ref{fig:cps} is a floating figure with four combinatorial proofs.

%%\figcpscondensed Fig.\,\ref{fig:cps-condensed} is a floating figure with the condensed forms of the four combinatorial proofs in Fig.\,\ref{fig:cps}.

@@

@@

\begin{figure*}
  $$
  \begin{tikzpicture}\cponeonformula
  \end{tikzpicture}
  \qquad
  \begin{tikzpicture}\cptwoonformula
  \end{tikzpicture}
  \qquad
  \begin{tikzpicture}\cpthreeonformula
  \end{tikzpicture}
  \qquad
  \begin{tikzpicture}\cpfouronformula
  \end{tikzpicture}
  $$
  \vskip-2ex
  \caption{Four combinatorial proofs, each shown
    above the formula proved.  Here $x$ and $y$ are variables, $f$ is
    a unary function symbol, $a$ and $b$ are constants (nullary
    function symbols), $p$ is a unary predicate symbol, and $q$ is a
    binary predicate symbol.}
  \label{fig:cps}
\end{figure*}

\begin{figure*}
  $$
  \begin{tikzpicture}
    \cponeinline
  \end{tikzpicture}
  \qquad\;
  \begin{tikzpicture}
    \cptwoinline
  \end{tikzpicture}
  \qquad\;
  \begin{tikzpicture}
    \cpthreeinline
  \end{tikzpicture}
  \qquad\;
  \begin{tikzpicture}
    \cpfourinline
  \end{tikzpicture}
  $$
  \vskip-2ex
  \caption{Condensed forms of the four combinatorial proofs in Fig.\,\ref{fig:cps}.}
  \label{fig:cps-condensed}
  %\vspace{-4ex}
\end{figure*}

@@

To give some examples, we show in Figure~\ref{fig:cps} for examples of
FOCPs, each above the formula it proves.  The same FOCPs are shown in
Figure~\ref{fig:cps-condensed} in a more condensed notation.

@@

\begin{remark}
  In or definition of FOCP, we are slightly laxer than the original
  definition of~\cite{hughes:fopws}, as we allow for a variable
  renaming $\substof\phi$ which was forced to be the identity
  in~\cite{hughes:fopws}.
  %The reason for this relaxation in our case is that we want to cover a richer class of proofs, also allowing 
\end{remark}

@@

\begin{example}
  Below is an example of a decomposed derivation in $\FOKS$ of the
  formula $(\exists x. \cneg{p}(x)) \vlor (\forall y.(p(y) \vlan
  p(f(y))))$:
  \begin{equation*}
    \vlderivation{
      \vlin{\acD}{}{(\exists x. \cneg{p}x) \vlor (\forall y.(py \vlan pfy))}{
        \vlin{\mexD}{}{(\exists x. (\cneg{p}x \vlor \cneg{p}x) \vlor (\forall
          y.(py \vlan pfy))}{
          \vlin{\fequ}{}{((\exists x. \cneg{p}x) \vlor (\exists x.\cneg{p}x))
            \vlor (\forall y.(py \vlan pfy)}{
            \vlin{\exists}{}{\forall y.(((\exists x. \cneg{p}x) \vlor (\exists
              x.\cneg{p}x)) \vlor (py \vlan pfy))}{
              \vlin{\exists}{}{\forall y.((\cneg{p}y \vlor (\exists x.\cneg{p}x))
                \vlor (py \vlan pfy))}{
                \vlin{\fequ}{}{\forall y.((\cneg{p}y \vlor \cneg{p}fy) \vlor
                  (py \vlan pfy))}{
              \vlin{\sw}{}{\forall y.(\cneg{p}y \vlor ((py \vlan pfy)
                \vlor \cneg{p}fy))}{
                \vlin{\fequ}{}{\forall y.(\cneg{p}y \vlor (py \vlan (pfy
                  \vlor \cneg{p}fy)))}{
                  \vlin{\aiD}{}{\forall y.((\cneg{p}y \vlor py) \vlan (pfy
                    \vlor \cneg{p}fy))}{
                    \vlin{\aiD}{}{\forall y.((\cneg{p}y \vlor py) \vlan
                      \ttt)}{
                      \vlin{\ttt}{}{\forall y.(\ttt \vlan \ttt)}{
                        \vlin{\forall}{}{\forall y.\ttt}{
                          \vlhy{\ttt}}}}}}}}}}}}}}
  \end{equation*}
\end{example}
